---
title: "Dixie Effect on Economic Stability in Australia"

author: "Ben Gussen"

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This research project investigates the dynamic effects of a weaker United States Dollar (USD) on the Australian economy, with a focus on economic stability. By employing Bayesian Structural Vector Autoregression (SVAR) analysis, this study aims to elucidate the transmission mechanisms of USD fluctuations through three channels: GDP growth, inflation, and unemployment. The findings are expected to provide nuanced insights into the macroeconomic interdependencies between the USD and the Australian economy, offering valuable perspectives for policymakers and economic analysts.
>
> **Keywords.** Bayesian SVAR, USD, Australian economy, inflation, economic stability

# Introduction

This document presents the design and implementation of a Bayesian Structural Vector Autoregression (SVAR) model. Our focus is on understanding the impact of the Trade Weighted Index (TWI) of the USD on Australian economic indicators: unemployment, inflation, and GDP rate.

# Research Proposal Structure

## The Question, Objective, and Motivation

**Objective:** To examine the dynamic effects of a weakening USD on the Australian economy, focusing on changes to the GDP, inflation and unemployment.

**Research Question:** How does a weakening United States Dollar (USD) influence overall economic stability in Australia?

**Motivation:** The relationship between the USD and the Australian economy is crucial, given the extensive trade links and financial interactions between the two nations. In light of recent global financial uncertainties, understanding this interplay is essential for crafting informed economic policies and strategies. This research endeavors to dissect the complexities of this economic relationship, aiming to provide insights that could inform both policymakers and market participants.

#### The US dollar trade weighted index or the USD/AUD exchange rate?

@Chen2004 offers insights into the dynamics of a small and open economy, discovering evidence from New Zealand that movements in exchange rates impact the value of firms. We argue that this impact could be wider, in partucluar, impacting macroeconomic indicators. The ability to ascertain the significance of such potential impact would rquire looking at the glboal relative strength of key currencies, in particular, in the case of Australia, in relation to the greenback. 

The U.S. Dollar Index (USDX, DXY, DX, or informally, the "Dixie") is an index measuring the value of the United States dollar relative to a basket of foreign currencies, often those of major U.S. trade partners [@Chen2024]. This index increases when the U.S. dollar gains strength compared to other currencies. Managed by ICE (Intercontinental Exchange, Inc.), it's a geometric mean weighted by:

- Euro (EUR) at 57.6%
- Japanese yen (JPY) at 13.6%
- Pound sterling (GBP) at 11.9%
- Canadian dollar (CAD) at 9.1%
- Swedish krona (SEK) at 4.2%
- Swiss franc (CHF) at 3.6%

Originating in March 1973 post-Bretton Woods, it started at 100.000, reaching a high of 164.7200 in February 1985 and a low of 70.698 on March 16, 2008. The index's composition changed once, in 1999, with the introduction of the euro, replacing several European currencies. There have been calls for updating the "basket" to reflect current major U.S. trading partners like China and Brazil, as Sweden and Switzerland have lesser trade volumes with the U.S.

The USD Index is influenced by major economic events and policies, demonstrated through historical data from 1967 to the present, showing fluctuations due to various global and domestic situations like oil crises, presidential elections, and economic policies. The index is also available through futures traded on the ICE platform and is calculated using a specific formula involving exchange rates against the aforementioned currencies.

The index is a vital tool for traders, policymakers, and economists as it offers insights into the relative strength of the U.S. dollar and, by extension, the relative health of the U.S. economy compared to its major trading partners. It reflects the impact of U.S. economic conditions, policies, and global events on the value of the dollar and is a crucial benchmark in the financial markets.

The US Trade Weighted Index and the USD/AUD exchange rate are both measures related to the value of the US dollar, but they provide different information. The USD Index is a broad measure of the value of the US dollar relative to a basket of currencies of its major trading partners. It is a weighted average, where currencies of more important trading partners have a greater influence on the index value. The weights are adjusted periodically to reflect changes in trading patterns. The USD Index focuses on the competitiveness of US goods and services in international trade. A higher USD Index indicates a stronger US dollar compared to the basket of currencies. 

On the other hand, the USD/AUD exchange rate is the price of one US dollar in terms of Australian dollars. This rate forcuses on the direct exchange value between the US dollar and the Australian dollar. A higher USD/AUD rate means it takes more Australian dollars to buy one US dollar (stronger USD).

In terms of the differences between the USD Index and the USD/AUD rate, the scope of the former is much larger than the scope of the latter, thus giving a better picture of the overall strength or weakness of the US dollar. USD Index considers a basket of currencies, while USD/AUD is just one currency pair. The USD Index focus reflects overall trade competitiveness, while USD/AUD is a specific exchange rate. The USD Index provides a broader picture, while USD/AUD is a more granular data point. Imagine the USD Index as a stock market index for the US dollar, reflecting its performance against a group of major currencies. The USD/AUD exchange rate is like the price of a single stock in that index.

In order to ensure capturing an international perspective of the US dollar position, it is more appropriate to use the USD Index instead of the USD/AUD exchage rate. This research project will therefore opt to employ the Dixie to check its effect on the economic health of the Australian economy. 

#### What is economic stability?

Economic stability describes a healthy condition of a nation's economy, characterized by:

* **Low and stable inflation:** Prices rise slowly and predictably, allowing people to plan for the future without fearing excessive erosion of purchasing power.
* **Moderate unemployment rate:** Enough jobs are available to employ most people seeking work, but not so few that wages become inflated due to a limited labor pool. 
* **Minimal fluctuations:** The economy experiences steady growth with minimal booms and busts. This translates to small variations in output growth.

Governments and central banks use various tools like fiscal policy (taxation and spending) and monetary policy (interest rates) to try and achieve this economic stability. It is a delicate balancing act, and achieving all these goals simultaneously can be challenging.

In simpler terms, economic stability reflects a predictable and healthy economic environment.  This stability fosters business confidence, consumer spending, and overall economic well-being. Measuring inflation, unemployment rate, and GDP growth is a proxy for measuring economic stability.

#### Effect of a weak US dollar on Australian inflation

The effect of a weaker US dollar on inflation in Australia involves several mechanisms, primarily through trade and investment channels. A weaker US dollar makes US imports cheaper in Australian terms. This can reduce the price of goods and services Australia imports from the US, potentially lowering the inflation rate, especially for products directly imported from the US or priced in US dollars. Also, since Australia is a significant exporter, a weaker US dollar can increase the local currency value of its US dollar-denominated exports. This leads to higher income for Australian exporters and can boost the economy. However, this can also contribute to cost-push inflation if the increased demand raises domestic prices. A weaker US dollar can enhance the competitiveness of Australian exports by making them more affordable internationally. This could increase export revenues, boost economic activity, and potentially lead to higher demand-pull inflation in Australia. As to investment Flows, a depreciating US dollar might redirect investment flows away from the US and towards other economies, including Australia. Increased investment can stimulate the Australian economy but also contribute to inflationary pressures if it leads to increased spending. In summary, a weaker US dollar can have inflationary effects by making imports cheaper, increasing the local currency earnings from exports, and improving the competitiveness of Australian goods in the global market. However, the actual impact on Australian inflation will depend on a variety of factors, including the overall trade balance, the elasticity of demand for imports and exports, monetary policy responses, and the state of the global and domestic economies. Notwithstanding, by measuring the historcial effect of the US Trade Weighted Index on domestic inflation, one can ascertain the aggregate effect of these other factors, and potentially, predict the future effect of a weakening of the US dollar on the Australian CPI rate. 

#### Effect of a weak US dollar on Australian unemployment 

A weaker US dollar generally translates to a stronger Australian dollar (AUD). With a stronger AUD, Australian goods and services become relatively cheaper on the global market. This can lead to increased demand for Australian exports, potentially boosting production and requiring more workers, which could lead to a decrease in unemployment. However, a stronger AUD can also make imports cheaper for Australian businesses. This can reduce production costs and potentially improve profit margins. However, it can also make domestically produced goods less competitive compared to cheaper imports, potentially leading to job losses in some sectors. The net effect on unemployment depends on the relative strengths of these two forces. If the export boost outweighs the import competition, the unemployment rate might decrease. However, the impact can be complex and depend on other economic factors. 

#### Effect of a weak US dollar on Australian GDP growth 

Similarly, A weaker US dollar can have both positive and negative effects on the Australian GDP growth rate. A weaker US dollar can make Australian exports relatively cheaper for US buyers, potentially boosting demand for Australian goods and services in the US market. This increase in exports could lead to higher GDP growth as Australian businesses expand to meet this demand. On the flip side, a weaker US dollar can make imports into Australia more expensive, which could lead to higher costs for businesses and consumers. If the increase in import prices outweighs the benefits from increased exports, it could dampen overall economic growth. In addition, Australia is a major exporter of commodities such as iron ore, coal, and natural gas. Commodity prices are often denominated in US dollars, so a weaker US dollar can lead to higher commodity prices in US dollar terms. This can be beneficial for Australian exporters and government revenues, contributing positively to GDP growth. Also, changes in currency values can influence investment flows between countries. A weaker US dollar might make Australian assets relatively more attractive to investors, leading to increased foreign investment inflows into Australia. This can stimulate economic activity and contribute to GDP growth. Overall, the net effect of a weaker US dollar on Australian GDP growth rate will depend on various factors including the responsiveness of exports and imports to changes in currency values, commodity price movements, and the dynamics of investment flows.

The transmission mechanisms through which the dynamics of the USD Index influence the economic stability indicators of the Australian economy, specifically inflation, unemployment, and GDP growth, are multifaceted and significant due to the interdependent nature of global trade and financial flows. Understanding these mechanisms is crucial for policymakers, businesses, and investors as they navigate the impacts of international monetary fluctuations.

1. **Trade Competitiveness and Inflation:** The USD Index represents the strength of the US dollar relative to a basket of currencies of its major trading partners, including Australia. A rise in the USD Index implies a relatively weaker Australian dollar, making Australian exports cheaper and more competitive on the global market. This increased competitiveness can boost Australia's export volumes, contributing positively to GDP growth. However, the flip side is that a weaker Australian dollar makes imports more expensive, translating into imported inflation. Imported goods, including essential items like oil, machinery, and technology, become pricier, pushing up the overall consumer price index (CPI) and thereby inflation rates within Australia. This imported inflation can erode purchasing power and increase the cost of living for Australians.

2. **Investment Flows and Unemployment:** Changes in the USD Index can influence international investment flows. A stronger US dollar (or a weaker Australian dollar) could divert investment from the US to Australian assets, seeking higher returns. This influx of capital can stimulate the Australian economy, potentially leading to job creation and lower unemployment rates. However, the volatility associated with fluctuating exchange rates may also lead to hesitancy among businesses to commit to new investments and expansions, possibly leading to a stagnation or increase in unemployment levels. Thus, the impact on unemployment can be dual-faceted, depending on other economic conditions and the response of businesses to currency movements.

3. **GDP Growth through Trade Balances:** The trade balance is a significant component of GDP calculation. As mentioned, a weaker Australian dollar (stemming from a higher USD Index) makes Australian goods more competitive abroad, potentially improving the trade balance as exports increase. However, this must be balanced against the cost of higher imports. If the increase in export revenue outweighs the increased costs of imports, the net effect could be positive for Australia's GDP growth. Conversely, if import costs rise substantially without a corresponding increase in exports, this could negatively impact GDP growth.

4. **Interest Rates and Monetary Policy:** The dynamics of the USD Index can indirectly affect Australia's monetary policy decisions, which in turn impact inflation and economic growth. For example, if the Reserve Bank of Australia (RBA) responds to a weaker Australian dollar and rising imported inflation by increasing interest rates, borrowing costs would rise. While this may help contain inflation, it could also dampen economic growth by making loans more expensive for consumers and businesses, potentially leading to reduced spending and investment.

5. **Consumer Confidence and Market Sentiments:** The strength of the US dollar and the corresponding weakness of the Australian dollar can influence consumer confidence and market sentiments. A depreciating domestic currency may lead to uncertainty and reduced consumer spending, particularly if inflation rises and purchasing power declines. This reduction in consumer demand can slow economic growth and increase unemployment.

6. **Global Economic Conditions and Commodity Prices:** Australia's economy is significantly influenced by global economic conditions and commodity prices, many of which are denominated in US dollars. Changes in the USD Index can affect commodity prices and, by extension, Australia's export revenues, given its status as a major commodity exporter. This relationship underscores how fluctuations in the USD Index can have far-reaching impacts on Australia's economic stability.

In conclusion, the dynamics of the US dollar index have substantial and varied transmission mechanisms affecting the Australian economy's inflation, unemployment, and GDP growth. These mechanisms underscore the interaction between exchange rates and economic stability, highlighting the importance of vigilant monetary and fiscal policies to mitigate adverse effects and capitalize on favorable movements.

## Data and Their Properties

In order to conduct a thorough analysis of the impact of USD fluctuations on the Australian economy, this study will utilize a set of time-series data that encapsulates three economic indicators:

- **Growth Rate in the Australian Gross Domestic Product:** Quarterly GDP figures representing the overall economic activity within Australia.

- **Unemployment Rate:** The benchmark interest rates set by the Reserve Bank of Australia, indicative of the cost of borrowing in the economy.

- **Inflation Rate:** Measured by the change in the Consumer Price Index (CPI), reflecting the changes in prices for goods and services.

These indicators will be analyzed against one proxy for the strength of the USD dollar as an interantional currency. 

- **USD Trade Weighted Index:** monthly index of US dollar major currency trade-weighted index. 


### Motivation for Data Choice:

The chosen variables provide a comprehensive view of the economic landscape necessary for understanding the dynamics between the USD and the Australian economy. Trade weighted indices directly influence the competitiveness of a country's goods and services abroad, impacting trade balances and, subsequently, GDP levels. Given that the United States is one of Australia's major trading partners, changes in the strength of the USD would be likely to impact the growth rate in Australia's GDP. Similarly, changes to the same index are likely to influence the inflation and unemployment rates, which are critical indicators of economic health, influencing purchasing power and consumer spending. Together, these data series allow for an in-depth analysis of international economic interactions and the stability of the Australian economy.

### Data Acquisition and Transformation:

The required data will be sourced from official Australian government databases, in particular the Reserve Bank of Australia. An initial look at the data indicates that they have different frequencies.

It is apposite to point out here that the project will conduct the analysis on data in the time window between 1980 and 2019. Data older than the first quarter of 1980 is likely to be too remote to be relevant to the current trade relations impacting on the USD and the Australain economy. Data points after the fourth quarter in 2019 were impacted by the COVID-19 crisis, and this is likely to distort  the analysis. 

Dealing with time series data at different frequencies within a Bayesian Structural Vector Autoregression (SVAR) model requires careful preprocessing. Mixed frequency data complicates modeling because the SVAR model requires observations at regular intervals. 

To decide on a common frequency for the  model, one needs to consider the research question and the nature of the data. If short-term dynamics are crucial, one should consider daily or monthly frequencies. For longer-term analysis, quarterly data might suffice. Given that the project is interested in identifying the effects of a weakening USD on the Australian economy, and given the time it would take for shocks to propagate through international trade, it would be reasonable to choose a longer frequency, such as a quarterly or an annual one, to allow for capturing the effect of a weakening USD on the Australian economy. However, because the window of analysis is reduced to between the years of 1980 and 2019, and to ensure there is enough data points to conduct a meaningful analysis, all four series will need to be converted into quarterly frequency, if they are not already in that format. 

For the purposes of this project, a daily series will be converted to a quarterly series by choosing the last day in the relevant month for each quarter.For example, for the first quarter, we use the last data point from March. For monthly series we use the data for the month corresponding to the end of each quarter (March, June, Sept, Dec).

Note, however, that one can also choose handling mixed frequencies in the analysis. Potentially, this is a more robust approach in terms of being faithful to the actual observed data. Techniques that are helpful here include: 

-	**Mixed Data Sampling (MIDAS)**: MIDAS regressions or other mixed-frequency approaches can be employed to incorporate variables of different frequencies directly into the analysis without needing to aggregate or interpolate all series to the same frequency.

-	**Temporal Hierarchy**: If using Bayesian methods, one might construct models separately for different frequencies and then combine insights hierarchically, acknowledging different data granularities.

For the initial stage of analysis, we will continue with the aligned quarterly data. Based on the outcomes, it could still be advantageous to re-assess the advantage of a mixed frequency approach. 

After aligning frequencies, one has to ensure that all series are stationary before including them in the SVAR model. Differencing, logging, or detrending might be necessary depending on the variable characteristics.

-	In a Bayesian framework, one has to incorporate prior knowledge about how different frequency data relates and affects each other.

-	One has to use Bayesian techniques to estimate the model parameters, acknowledging the frequency differences either in the data preparation stage or model structure, depending on your approach.

For our purposes in this initial proposal, the required data will be sourced from official Australian government databases using the `readrba` R packages. Based on the preliminary results from the project, it might become necessary to entertain a mixed frequency approach. 

The first series, the year-end Australian real GDP growth rate, is from the Reserve Bank of Australia (RBA) output and labour statistical tables [@RBA2024a]. This is a quarterly series, and will not need to be treated for frequency alignment. 

```{r GDP growth rate}
#| echo: false
#| message: false
#| warning: false
#| label: fig-GDP
#| fig-cap: "ACF and PACF functions for the GDP growth rate"
library(readrba)
library(dplyr)
library(zoo)
library(lubridate)
library(tseries)
library(ggplot2)
library(urca)  # For conducting the ADF test
library(forecast) # For ACF and PACF plots 
library(gridExtra)
library(kableExtra)

# Downloading data from the internet
series <- read_rba(series_id = "GGDPCVGDPY")

#clean the series by removing missing values
series <- na.omit(series)

# Ensuring the date format is appropriate for quarterly data
series <- series %>% 
  mutate(date = as.yearqtr(date, format = "%Y-%q"))

# Filtering the dataset for the desired date range
GDP_rate <- series %>%
  filter(date >= as.yearqtr("1980 Q1") & date <= as.yearqtr("2019 Q4"))

# Set up the plotting area

par(mfrow=c(1,2))

# Plot ACF with lags limited to 10
acf(GDP_rate$value, main = "ACF for GDP Rate", lag.max = 10)
# Plot PACF with lags limited to 10
pacf(GDP_rate$value, main = "PACF for GDP Rate", lag.max = 10)
# Reset the plotting area back to default
par(mfrow=c(1, 1))
```

As seen in @fig-GDP, the PACF above, the optimal lag for conducting the augmented Dicky-Fuller (DF) test is 4. A summary of the results of this test can be found in the appendix. 

```{r}
#| echo: false
#| message: false
#| warning: false

# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(GDP_rate$value, type = "drift", lags = 4, selectlags = "AIC")
```

Note that the augmented DF test suggests that the series is stationary at the 95% level. There is no need to modify the series. From the DF test results, its is safe to conclude that the GDP_rate series is stationary. 

```{r}
#| echo: false
#| message: false
#| warning: false

# If the series is non-stationary, take the first difference and adjust dates
if (adf_test_result@teststat[1] > adf_test_result@cval[1]) {  # Check if non-stationary
  GDP_rate$value <- diff(GDP_rate$value)
  # Adjust dates accordingly
  GDP_rate$date <- GDP_rate$date[-1]
}

# Determine plot title based on whether the series has been differenced
if (is.null(attr(GDP_rate$value, "diff"))) {  # Check if series has been differenced
  plot_title <- "Quarterly GDP Rate (1980-2019)"
} else {
  plot_title <- "Differenced Quarterly GDP Rate (1980-2019)"
}

```

The second series, the year-end change in Australian inflation rate, is from the Reserve Bank of Australia (RBA) inflation and inflation expectations statistical tables (@RBA2024b). This series has been converted from a monthly series into a quarterly series by chosing the relevant months for the beginning of each quarter, starting from March and ending with December. 

```{r CPI rate}
#| echo: false 
#| message: false
#| warning: false
#| label: fig-CPI
#| fig-cap: "ACF and PACF functions for the CPI rate"
library(ggplot2)
library(dplyr)
library(readrba)
library(urca)  # For conducting the ADF test

## Downloading CPI data from the internet
series <- read_rba(series_id = "GCPIAGYP") 

#clean the series by removing missing values
series <- na.omit(series)

# Ensuring the date format is appropriate for quarterly data
series <- series %>% 
  mutate(date = as.yearqtr(date, format = "%Y-%q"))

# Ensuring that date is recognized as a quarterly object and summarizing to quarterly values
series_quarterly <- series %>% 
  mutate(quarter = as.yearqtr(date)) %>%
  group_by(quarter) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Filter for the correct range from Q1 1980 to Q4 2019
CPI_rate <- series_quarterly %>%
  filter(quarter >= as.yearqtr("1980 Q1") & quarter <= as.yearqtr("2019 Q4"))

# Set up the plotting area
par(mfrow=c(1, 2))

# Plot ACF with lags limited to 10
acf(CPI_rate$value, main = "ACF for CPI Rate", lag.max = 10)

# Plot PACF with lags limited to 10
pacf(CPI_rate$value, main = "PACF for CPI Rate", lag.max = 10)

# Reset the plotting area back to default
par(mfrow=c(1,1))

# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(CPI_rate$value, type = "drift", lags = 4, selectlags = "AIC")


# Difference the series if necessary
if (adf_test_result@teststat[1] > adf_test_result@cval[1, "5pct"]) {  
  CPI_rate <- CPI_rate %>%
    mutate(diff_value = c(NA, diff(value)))  # Prepend NA to maintain the length
} else {
  message("The series is already stationary.")
}

# Remove NA values from the differenced series
CPI_rate <- na.omit(CPI_rate)


```

From @fig-CPI, an augmented DF test was conducted with 3 lags. Note that the augmented DF test (see appendix) suggests that the series is not stationary at the 95% confidence level. There was a need to take the difference of the series.

The third series, the Australian unemployment rate, is from the Reserve Bank of Australia (RBA) output and labour Statistical tables [@RBA2024c]. This is  a montly series. The series was converted into a quaterly series by chosing the relevant months for the beginning of each quarter, starting from March and ending with December. 

```{r unemp_rate}
#| echo: false 
#| message: false
#| warning: false
#| label: fig-unemp
#| fig-cap: "ACF and PACF functions for the unemployment rate"
library(ggplot2)
library(dplyr)
library(readrba)
library(zoo)  # For as.yearqtr and aggregation to quarters
library(urca)  # For conducting the ADF test

## Downloading unemployment data from the internet
series <- read_rba(series_id = "GLFSURSA") 

#clean the series by removing missing values
series <- na.omit(series)

# Convert the date format from dd/mm/yyyy to Date object
series <- series %>% 
  mutate(date = as.Date(as.character(date), format = "%Y-%m-%d")) 

# Ensuring that date is recognized as a quarterly object and summarizing to quarterly values
series_quarterly <- series %>% 
  mutate(quarter = as.yearqtr(date)) %>%
  group_by(quarter) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Filter the series to start from the first quarter of 1980 and end in the last quarter of 2019
unemp_rate <- series_quarterly %>%
  filter(quarter >= as.yearqtr("1980 Q1") & quarter <= as.yearqtr("2019 Q4"))

# Set up the plotting area
par(mfrow=c(1, 2))

# Plot ACF with lags limited to 10
acf(unemp_rate$value, main = "ACF for Unemployment Rate", lag.max = 10)

# Plot PACF with lags limited to 10
pacf(unemp_rate$value, main = "PACF for Unemployment Rate", lag.max = 10)

# Reset the plotting area back to default
par(mfrow=c(1, 1))

# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(unemp_rate$value, type = "drift", lags = 4, selectlags = "AIC")


# Check stationarity and apply differencing if necessary
if(adf_test_result@teststat[1] <= adf_test_result@cval[1, "5pct"]) {  # Adjust based on desired significance level
  # Plotting the data without differencing
  ggplot(unemp_rate, aes(x = quarter, y = value)) +
    geom_line() +
    theme_minimal() +
    labs(title = "Quarterly Unemployment Rate (1980-2019)", x = "Quarter", y = "Unemployment Rate (%)")
} else {
  # Plotting the differenced data
  unemp_rate <- unemp_rate %>%
    mutate(diff_value = c(NA, diff(value)))  # Prepend NA to maintain the length
}
```

@fig-unemp above explains that a lag 4 should be used for the augmented DF test. Note that this test (see Appendix) suggests that the series is not stationary at the 95% confidence level. We therefore use the differenced series.

The fourth series, the USD Index, is also from the Reserve Bank of Australia (RBA) exchange rates statistical tables [@RBA2024d]. This is  a montly series. The series was converted into a quaterly series by chosing the relevant months for the beginning of each quarter, starting from March and ending with December.

```{r USD Index}
#| echo: false
#| message: false
#| warning: false
#| label: fig-Index
#| fig-cap: "ACF and PACF functions for the USD Trade Weighted Index"
library(ggplot2)
library(dplyr)
library(readrba)
library(urca)  # For conducting the ADF test
library(zoo)  # For as.yearqtr

# Downloading USD Trade Weighted Index data from the internet (March 1973 = 100)

# Reading and preparing the USD TWI data
series <- read_rba(series_id = "FUSXRTWI") 

#clean the series by removing missing values
series <- na.omit(series)

# Convert the date format from dd/mm/yyyy to Date object
series <- series %>% 
  mutate(date = as.Date(as.character(date), format = "%Y-%m-%d")) 

# Ensuring that date is recognized as a quarterly object and summarizing to quarterly values
series_quarterly <- series %>% 
  mutate(quarter = as.yearqtr(date)) %>%
  group_by(quarter) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Filter the series to start from the first quarter of 1980 and end in the last quarter of 2019
USD_Index <- series_quarterly %>%
  filter(quarter >= as.yearqtr("1980 Q1") & quarter <= as.yearqtr("2019 Q4"))

# Set up the plotting area
par(mfrow=c(1, 2))

# Plot ACF with lags limited to 10
acf(unemp_rate$value, main = "ACF for USD Index", lag.max = 10)

# Plot PACF with lags limited to 10
pacf(unemp_rate$value, main = "PACF for USD Index", lag.max = 10)

# Reset the plotting area back to default
par(mfrow=c(1, 1))

# Augmented Dickey-Fuller test to check for stationarity
adf_test_result <- ur.df(USD_Index$value, type = "drift", lags = 4, selectlags = "AIC")

# Difference the series if necessary
if (adf_test_result@teststat[1] > adf_test_result@cval[1, "5pct"]) {  
  USD_Index <- USD_Index %>%
    mutate(diff_value = c(NA, diff(value)))  # Prepend NA to maintain the length
} else {
  message("The series is already stationary.")
}

# Since the first value is now NA after differencing, remove it for plotting
USD_Index <- na.omit(USD_Index)

```

@fig-Index above indicates that a lag 4 should be used for the augmented DF test. The results suggest that the series is not stationary at the 95% confidence level. We therefore use the differenced series.

The four treated series for the proposed model are in  @fig-combined below. All these series are now reflecting change (rates) in the underlying economic indicator. 

```{r plotting the four series}
#| echo: false
#| message: false
#| warning: false
#| label: fig-combined
#| fig-cap: "Treated series plots"
library(ggplot2)
library(gridExtra)

# GDP_rate, CPI_rate, unemp_rate, and USD_Index are already prepared and available

p1 <- ggplot(GDP_rate, aes(x = date, y = value)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Quarterly GDP Rate (1980-2019)", x = "Quarter", y = "GDP Growth Rate (%)")

p2 <- ggplot(CPI_rate, aes(x = quarter, y = diff_value)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Differenced Quarterly CPI Rate (1980-2019)", x = "Quarter", y = "CPI Rate (%)")

p3 <- ggplot(unemp_rate, aes(x = quarter, y = diff_value)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Differenced Quarterly Unemployment Rate (1980-2019)", x = "Quarter", y = "Differenced Unemployment Rate (%)")

p4 <- ggplot(USD_Index, aes(x = quarter, y = diff_value)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Differenced Quarterly USD Index (1980-2019)", x = "Quarter", y = "USD Trade Weighted Index (%)")

# Adjust the title size using theme() function
p1 <- p1 + theme(plot.title = element_text(size = 9))
p2 <- p2 + theme(plot.title = element_text(size = 9))
p3 <- p3 + theme(plot.title = element_text(size = 9))
p4 <- p4 + theme(plot.title = element_text(size = 9))

# Combine the plots
grid.arrange(p1, p2, p3, p4, ncol = 2)

```

The above selection and treatment of the data is pivotal to answering the research question: How does a weakening United States Dollar (USD) influence overall economic stability in Australia? By focusing on key economic indicators like the GDP growth rate, unemployment rate, and inflation rate, and comparing these against the USD Trade Weighted Index (TWI), the study strategically addresses the multifaceted impact of USD fluctuations on the Australian economy. The chosen data encapsulate the essential aspects of economic stability, allowing for a nuanced analysis of the interdependencies between the USD's value and Australian economic health.

This approach is vital for constructing a comprehensive Bayesian Structural Vector Autoregression (SVAR) model. The SVAR model, designed to elucidate the transmission mechanisms of USD fluctuations, relies on accurate, time-aligned, and contextually relevant data to provide meaningful insights. By ensuring the data are stationary and appropriately preprocessed for the chosen analysis timeframe (1980-2019), excluding periods of atypical economic disruption like the COVID-19 crisis, the document ensures that the model's outputs are reflective of standard economic interactions, enhancing the reliability of the findings.

Furthermore, the data treatment, including the conversion of series to quarterly frequencies and addressing stationarity, directly informs the model equations by ensuring that the inputs reflect genuine economic trends devoid of seasonal or non-stationary noise. This data preparation not only strengthens the study’s methodological framework but also ensures that the SVAR model's findings will offer actionable insights, enabling policymakers and analysts to base decisions on solid empirical evidence, thereby underlining the importance of investigating the specified problem.

Next, we employ a Baesian Structural Vector Autoregression (SVAR) model to understand the dynamic relationships among various economic indicators influenced by fluctuations in the United States Dollar (USD) against the Australian Dollar (AUD) [@wozniakBsvarsBayesianEstimation2022]. In particular, we follow @Bhuiyan2012 who introduces a Bayesian structural VAR model tailored for Canada, focusing on evaluating the impact of monetary policy shocks, utilizing the overnight rate target as the main policy tool. However, our model differs in that it does not directly account for any fiscal and monetary policy,  rather it evaluates the potential of using the US Index to predict the stability of the Australian economy. 

# Model Specification

This is a 4x1 vector of endogenous variables at time t, where:

- $USD\_Index$ is the Trade Weighted Index of the US Dollar.
- $Unemployment\_rate$ is the unemployment rate.
- $Inflation\_rate$ is the inflation rate.
- $GDP\_growth$ is the growth rate of GDP.


The SVAR model structure:

- $A$ is a 4X4 matrix representing the contemporaneous relationships between the variables.
- $B_i$ (for i = 0,1,...,p) are 4x4 matrices representing the lagged effects (where p is the lag order).
- $\epsilon_t$ is a 4x1 vector of errors.

$$ 
Y_t =
\begin{bmatrix}
    USD\_Index \\
    Unemployment\_rate \\
    Inflation\_rate \\
    GDP\_growth
\end{bmatrix}
$$

The Bayesian Structural VAR model can then be written as: 

$$
AY_t = 
B_0 + B_1Y_{t-1} + ... + B_pY_{t-p} + \epsilon_t
$$

#### Model Notations:

$Yt$​: This represents the vector of endogenous variables at time tt. In this context, the vector consists of the Trade Weighted Index of the US Dollar (USD_Index), the unemployment rate, the inflation rate, and the GDP growth rate. These variables are chosen because they represent key indicators of economic stability.

$A$: This is a matrix representing the contemporaneous relationships between the variables in $Yt$. In an SVAR model, this matrix helps to understand how shocks to one variable (e.g., USD_Index) can contemporaneously affect other variables (e.g., unemployment, inflation, GDP growth).

$Bi$ (for i=0,1,...,pi=0,1,...,p): These are matrices representing the lagged effects of the variables on each other. The subscript ii represents the lag order, showing how past values (lags) of each variable in $Yt$​ influence the current values. In time series analysis, this captures dynamics such as how past economic conditions influence current outcomes.

$ϵt$​: This is the vector of errors or shocks at time tt. In the SVAR framework, these are considered unobservable random shocks that affect the endogenous variables. In Bayesian SVAR models, these shocks are also subjected to prior distributions reflecting our beliefs or assumptions before observing the data.

#### Using the Model to Answer the Research Question:

The model will be used to analyze how fluctuations in the US Trade Weighted Index (representing a weakening or strengthening USD) impact the Australian economy's stability, specifically through GDP growth, inflation, and unemployment rates. By examining the contemporaneous and lagged relationships between these variables, we can understand the transmission mechanisms of US dollar fluctuations.

To address the research objectives, we will assess the dynamic effects of the USD_Index on Australia's economic indicators. This involves looking at the impulse response functions (IRFs) and variance decompositions derived from the SVAR model. The IRFs will show how a shock to the US dollar index impacts Australian GDP growth, inflation, and unemployment over time. This directly speaks to the study's aim to elucidate the transmission mechanisms of such fluctuations.

In addition, to enhance our understanding of the dynamic relationships among economic indicators and the USD Index, we will incorporate an analysis of Granger causality (see @Kirchgässner&Wolters2007, pp. 93-123). This addition aims to rigorously test the directional influence between the USD Index and key Australian economic metrics: inflation, unemployment, and GDP growth. To implement this, we propose extending the Bayesian Structural Vector Autoregression (SVAR) model to include Granger causality tests. These tests will determine whether past values of the USD Index can predict future values of Australian inflation, unemployment, and GDP growth, beyond the predictions based on past values of these Australian indicators themselves.

This analysis will be conducted using a stepwise approach:

Lag Selection: Optimal lags for the variables will be determined based on information criteria such as AIC or BIC within the Bayesian framework.

Model Estimation: The extended SVAR model, incorporating the selected lags, will be estimated to analyze the interactions between the USD Index and Australian economic indicators.

Granger Causality Tests: We will perform Bayesian Granger causality tests within the context of the SVAR model to investigate if the USD Index has a predictive power over Australian inflation, unemployment, and GDP growth rates.

The addition of Granger causality analysis will provide a more nuanced understanding of how changes in the USD Index might precede alterations in Australia’s economic conditions. This could be particularly insightful for policymakers, providing evidence on whether movements in the USD Index could serve as a leading indicator for shifts in economic stability within Australia.

By integrating this causality testing into our model, we aim to offer more comprehensive insights into the interdependencies between international currency movements and domestic economic health. This approach will help in clarifying the temporal precedence and potential predictive relationships, thereby enriching the findings and implications of our research on the impacts of the USD's strength on Australian economic stability.

#### Estimation Outputs for Interpreting the Research Question:

Posterior Distributions: These reflect the updated beliefs about the model's parameters after considering the data. We would examine the posterior distributions of the SVAR model parameters, particularly focusing on the coefficients that measure the impact of USD_Index changes on the economic indicators.

Impulse Response Functions (IRFs): IRFs trace the effects of a one-time shock to one of the innovations on the current and future values of the endogenous variables. By analyzing the IRFs, we can interpret how an unexpected change in the US dollar's value affects Australian economic conditions over time.

Variance Decompositions: These help us understand the proportion of the forecast variance of each endogenous variable that can be attributed to shocks to each variable in the model, including shocks to the USD_Index. This is crucial for assessing the significance of US dollar fluctuations compared to other domestic factors.

Credibility Intervals: Unlike classical confidence intervals, credibility intervals in Bayesian analysis offer a probability statement about the parameter values. Examining these intervals for the effects of interest will provide insights into the certainty of our estimates.

In terms of economic context, understanding the relationship between the US dollar index and Australian economic indicators is vital for policymakers and economic analysts, especially considering the significant trade and financial ties between the two countries. The model helps address the research objectives by providing a structured way to quantify these relationships, thereby informing decisions that could enhance economic stability in response to international currency movements.


## Appendix: Augmented Dicky-Fuller Tests 

### Augmented Dickey-Fuller test results for GDP growth 

```{r}
#| echo: false
#| message: false
#| warning: false
# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(GDP_rate$value, type = "drift", lags = 4, selectlags = "AIC")

summary(adf_test_result)  # Outputs the result of the test
```

### Augmented Dicky-Fuller test results for the CPI rate

```{r}
#| echo: false
#| message: false
#| warning: false
# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(CPI_rate$value, type = "drift", lags = 4, selectlags = "AIC")
summary(adf_test_result)  # Outputs the result of the test
```

### Augmented Dicky-Fuller test results for the unemployment rate

```{r}
#| echo: false
#| message: false
#| warning: false
# Conduct the ADF test to check for stationarity
adf_test_result <- ur.df(unemp_rate$value, type = "drift", lags = 4, selectlags = "AIC")
summary(adf_test_result)  # Outputs the result of the test
```

### Augmented Dicky-Fuller test results for the USD Index

```{r}
#| echo: false
#| message: false
#| warning: false
# Augmented Dickey-Fuller test to check for stationarity
adf_test_result <- ur.df(USD_Index$value, type = "drift", lags = 4, selectlags = "AIC")
summary(adf_test_result)  # Outputs the result of the test
```

# References{.unnumbered}



